Age-related changes in plant growth and traits

### 2015-04-30 results_growth.Rmd
more cleaning up of results

@@

### Species with larger seeds have higher growth rates  
- The same basic relationship exists across all ages, so combined all data here  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ seed_size,data=subset(SummarySppAge,age>.5),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="seed size (mg)",ylab="growth investment (mg per year)",main="Effect of seed size on growth rates")
legend("topleft",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ seed_size, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

### Species with greater maximum height have higher growth rates
- For younger plants there is no relationship with *SPECIES* overall maximum H  
- There is a positive correlation with the maximum height the species reaches at a given age -i.e. fast growing young plants are tall when young.  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age<3),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.1,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==2.4), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")

@@

### Species with higher wood density have lower growth rates  
- Adult plants strongly show this pattern
- However there is no relationship in younger plants  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==5), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomleft",col=col.age, labels.age, pch=16, cex=1, bty="n")

@@

>>>>

###Shifts in leafiness with age  
- Individual shoots looked decidedly "less leafy" in many older plants  
- This led me to wonder how leaf lifespan - or some other variable - was shifting with age   
- The number of leaves along a leader is overall uneffected by age  
- However, looking at individual species: BOLE, GRBU, GRSP, HEPU, PILI, PUTU have declining leaf counts along a shoot with age  
- And PELA, PEPU show a significant increase in leaf counts along a single shoot with age

```{r, echo=FALSE,message=FALSE}
plot(lvs_end_total~age,SummaryInd,log="xy", xlab="age",ylab="final leaf count along shoot",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1,bty="n")

yvar <- "lvs_end_total"
xvar <- "age"

###Leaf weight (relative to total weight) also declines with age  
- These numbers are currently incomplete - waiting for full set of "leaf weight" numbers to finalize this

```{r, echo=FALSE,message=FALSE}
plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1, bty="n")

plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.spp[as.factor(species)], pch=16, cex=0.8)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=.8, bty="n")
```

###Shifts in shoot extension with age
- The decrease in leaf number along a shoot is largely caused by declining shoot extension, not shifts in leaf lifespan  
- BAER, BOLE, EPMI, GRBU, GRSP, HATE, HEPU, PHPH, PILI, and PUTU all show significant declines in shoot extension with age  
- For many of these species, the shift is between the "sapling" and "adult" stages - i.e. when they begin reproducing  
- PELA stands out by showing increasing shoot extension with age - because it grows so slowly in the beginning

```{r, echo=FALSE,message=FALSE}
plot(new_length~age,SummaryInd,log="xy", xlab="age",ylab="shoot extension (mm)",col=col.spp[as.factor(species)])
legend(12,200,col=col.spp, labels.spp, pch=16, cex=.8,bty="n")

yvar <- "new_length"
xvar <- "age"

@@

### Species with larger seeds have higher growth rates  
- The same basic relationship exists across all ages, so combined all data here  
```{r, echo=FALSE,message=FALSE}
plot(GrowthInv_mean ~ seed_size,data=subset(SummarySppAge,age>.5),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="seed size (mg)",ylab="growth investment (mg per year)",main="Effect of seed size on growth rates")
legend("topleft",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ seed_size, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

### Species with greater maximum height have higher growth rates
- For younger plants there is no relationship with *SPECIES* overall maximum H  
- There is a positive correlation with the maximum height the species reaches at a given age -i.e. fast growing young plants are tall when young.  
```{r, echo=FALSE,message=FALSE}
plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age<3),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.1,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==2.4), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")

@@

### Species with higher wood density have lower growth rates  
- Adult plants strongly show this pattern
- However there is no relationship in younger plants  
```{r, echo=FALSE,message=FALSE}
plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==5), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomleft",col=col.age, labels.age, pch=16, cex=1, bty="n")

@@

### Species with higher LMA have higher growth rates  
- This pattern is true for adult plants  
- But there is no relationship in seedlings (age=1.4) or saplings (age=2.4)  
- This is not predicted by Daniel's model or Anais' review; re-run with RGR
```{r, echo=FALSE,message=FALSE}
plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==1.4),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Seedlings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")
mod <- lm(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4))
summary(mod)

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==2.4),log="xy",col="pink",pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Saplings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")
mod <- lm(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4))
summary(mod)

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Adults")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age>3), labels=species, pos=1, offset=-1.1, cex=0.6,col="blue")
legend("bottomright",col=col.age,labels.age,pch=16, bty="n")
mod <- lm(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age>3))
summary(mod)
```

###Looking at LMA relative to RGR
- There are no significant relationships at any age when RGR is used instead of actual growth investment 
```{r, echo=FALSE,message=FALSE}
plot(RGR_mean ~ LMA_mean,data=subset(SummarySppAge,age==1.4),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="RGR=",main="Seedlings")
text(RGR_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")
mod <- lm(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4))
summary(mod)

plot(RGR_mean ~ LMA_mean,data=subset(SummarySppAge,age==2.4),log="xy",col="pink",pch=16,cex=1.2,xlab="LMA",ylab="RGR",main="Saplings")
text(RGR_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")
mod <- lm(RGR_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4))
summary(mod)

plot(RGR_mean ~ LMA_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="RGR",main="Adults")
text(RGR_mean ~ LMA_mean, data=subset(SummarySppAge,age>3), labels=species, pos=1, offset=-1.1, cex=0.6,col="blue")
legend("bottomright",col=col.age,labels.age,pch=16, bty="n")
mod <- lm(RGR_mean ~ LMA_mean, data=subset(SummarySppAge,age>3))
summary(mod)
```

###Leaf lifespan vs growth investment
- There is no relationship between leaf lifespan and growth investment for any age group
- And just barely a significant relationship between RGR and leaflifespan for adult plants
```{r, echo=FALSE,message=FALSE}
plot(GrowthInv ~ LL_death,data=subset(SummaryInd,age==1.4),log="xy",col=col.spp[as.factor(species)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="growth investment (mg per year)",main="Seedlings")
legend(4.5,1600,col=col.spp,labels.spp,pch=16, bty="n",cex=.8)
mod <- lm(GrowthInv ~ LL_death, data=subset(SummaryInd,age==1.4))
summary(mod)

plot(GrowthInv ~ LL_death,data=subset(SummaryInd,age==2.4),log="xy",col=col.spp[as.factor(species)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="growth investment (mg per year)",main="Saplings")
legend("bottomright",col=col.spp,labels.spp,pch=16, bty="n",cex=.8)
mod <- lm(GrowthInv ~ LL_death, data=subset(SummaryInd,age==2.4))
summary(mod)

plot(GrowthInv ~ LL_death,data=subset(SummaryInd,age>3&LL_death<10),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="growth investment (mg per year)",main="Adults")
legend("bottomright",col=col.age,labels.age,pch=16, bty="n")
mod <- lm(GrowthInv ~ LL_death*species, data=subset(SummaryInd,age>3))
anova(mod)
```

```{r, echo=FALSE,message=FALSE}
plot(RGR ~ LL_death,data=subset(SummaryInd,age==1.4),log="xy",col=col.spp[as.factor(species)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="RGR",main="Seedlings")
legend(4.5,1600,col=col.spp,labels.spp,pch=16, bty="n",cex=.8)
mod <- lm(RGR ~ LL_death, data=subset(SummaryInd,age==1.4))
summary(mod)

plot(RGR ~ LL_death,data=subset(SummaryInd,age==2.4),log="xy",col=col.spp[as.factor(species)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="RGR",main="Saplings")
legend("bottomright",col=col.spp,labels.spp,pch=16, bty="n",cex=.8)
mod <- lm(RGR ~ LL_death, data=subset(SummaryInd,age==2.4))
summary(mod)

plot(RGR ~ LL_death,data=subset(SummaryInd,age>3&LL_death<10&RGR>0.04),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="leaf lifespan",ylab="RGR",main="Adults")
legend("bottomright",col=col.age,labels.age,pch=16, bty="n")
mod <- lm(RGR ~ LL_death*species, data=subset(SummaryInd,age>3&LL_death<10&RGR>0.04))
anova(mod)
```

### 2015-04-30 results_leaf_lifespan.rmd
more cleaning up of results

#Starting with the basics: Are leaf lifespan and LMA correlated?
Plotting species*site means
Figures shown using all three measures of leaf lifespan
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge)

@@

#Same plots as above, but omitting HATE, which has much higher LMA than any other species due to its round "leaves"
Using leaf deaths gives the poorest fit...
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge, species!="HATE")

@@

>>>>

###Starting with the basics: Are leaf lifespan and LMA correlated?  
- Plotting species*site means  
- Figures shown using all three measures of leaf lifespan  
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge)

@@

###Same plots as above, but omitting HATE, which has much higher LMA than any other species due to its round "leaves"  
- Using leaf deaths gives the poorest fit...
```{r, echo=FALSE,message=FALSE}
yvar <- "LMA_mean"
xvar <- "LL_death_mean"
data <- filter(SummarySppAge, species!="HATE")

@@

plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=15)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1,bty="n")
